Multi-domain, multilingual translation with latent multi-task group dropout

### 2021-11-16 2022/acl/acl_latex.tex
quick check of the appendix

\bibliography{bibliography}
\bibliographystyle{acl_natbib}
\appendix
\section{Appendix A}
\label{appendix:a}
This section explains how we calculate $\hat{m}_l^d(\tau)$ by solving the optimization problem \eqref{eq:soft-top-k} and then how to compute the gradient $\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}$.

First, to solve \eqref{eq:soft-top-k} we follow the same approach as in \cite{amos19lml,amos20differential} by applying the Karush–Kuhn–Tucker (KKT) conditions to \eqref{eq:soft-top-k}. The solution of \eqref{eq:soft-top-k} will have the following form

\begin{align}
\hat{m}_l^d(\tau) &= \sigma(\frac{g_l^d + \Phi_l^d + \bar{\nu}}{\tau}) \label{eq:soft-m}
\end{align}
in which $\sigma(.)$ is sigmoid function and $\bar{\nu}$ is the solution of the following equation
\begin{align}
\displaystyle{\mathop{\sum}_{i=1}^{n_p}} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) = k \label{eq:nu}
\end{align}

Because sigmoid is monotonic, the equation \eqref{eq:nu} has unique solution. Furthermore,  because of the smoothness of $g(\nu,\Phi_l^d) = \displaystyle{\mathop{\sum}_{i=1}^{n_p}} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau})$ w.r.t $\nu$ and $\Phi_l^d$, we could deduce the implicit differentiation of its solution $\bar{\nu}$ w.r.t $\Phi_l^d$ as below although the solution of \eqref{eq:nu} does not have explicit form.

@@

Because the differentiation of sigmoid has exact forms, $\frac{\partial g}{\partial \nu}$ and $\frac{\partial g}{\partial \Phi_l^d}$ also have exact form. Therefore, we do not need autograd to compute the implicit gradient $\frac{\partial \nu}{\partial \Phi_l^d}$. The gradient of $\hat{m}_l^d(\tau)$ w.r.t $\Phi_l^d$ is deduced as follows
\begin{align}
\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d} = \frac{\partial \hat{m}_l^d(\tau)}{\partial \nu} \times \frac{\partial \nu}{\partial \Phi_l^d}
\end{align}

In our algorithm, we solve \eqref{eq:nu} by binary search. The convergence of binary search is extremely fast and assured by the monotonicity of $g(\nu,\Phi_l^d)$. In our experiments, we set our search range to $[-100,100]$.

Finally, we need prove the limit $\lim_{\tau \rightarrow 0}\hat{m}_l^d(\tau) = \tilde{m}_l^d$. We suppose $g_l^d(i_1) + \Phi_l^d(i_1) > g_l^d(i_2) + \Phi_l^d(i_2) > \cdots > g_l^d(i_{n_p}) + \Phi_l^d(i_{n_p})$.

Because

\begin{align*}
\hspace{-3.em}
\lim_{\tau \rightarrow 0} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) = \begin{cases}
      1, & \text{if}\ \tau > - (g_l^d(i) + \Phi_l^d(i)), \\
      0, & \text{if}\ \tau < - (g_l^d(i) + \Phi_l^d(i)), \\
      \frac{1}{2} &\text{otherwise}\\
    \end{cases}
\end{align*}

@@

\begin{align*}
\displaystyle{\mathop{\sum}_{i=1}^{n_p}} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) = k
\end{align*}

there exist $\epsilon$ such that $\forall \tau < \epsilon$, the solution $\bar{\nu}$ of \eqref{eq:nu} is between $- (g_l^d(i_{k+1}) + \Phi_l^d(i_{k+1})) > \bar{\nu} > - (g_l^d(i_k) + \Phi_l^d(i_k))$. Furthermore, because sigmoid is monotonic,

@@

By taking the limit of both sides, we have following limits

\begin{align*}
\lim_{\tau \rightarrow 0} \hat{m}_l^d(\tau)(i_u) = \begin{cases}
      1, & \text{if}\ u > k \\
      0, & \text{if}\ u < k \\
    \end{cases}
\end{align*}

And, because $ \displaystyle{\mathop{\sum}_{u=1}^{n_p}}\hat{m}_l^d(\tau)(i_u) = k$, by taking the limit at both side, we will have $\lim_{\tau \rightarrow 0} \hat{m}_l^d(\tau)(i_k) = 1$. Finally, we have

@@

\section{Appendix B}
\label{appendix:b}
In this section, we give a simple proof of Inequality \ref{eq:entropy}. In fact, we only need to prove $\mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] \geqslant \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big]$. The proof is as follows
\begin{align*}
&\mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] = - \displaystyle{\mathop{\mathbb{E}}_{i_1,\cdots,i_k | \Phi_l^d}} \big[ \text{log}P(i_1,\cdots,i_k | \Phi_l^d) \big] \\
&= -\displaystyle{\mathop{\mathbb{E}}_{i_1,\cdots,i_k | \Phi_l^d}} \big[ \displaystyle{\sum_{j=2}^k}\text{log}P(i_j | i_1,\cdots,j_{j-1},\Phi_l^d) +  \text{log} P(i_1 | \Phi_l^d) \big] \\
&\geqslant -\displaystyle{\mathop{\mathbb{E}}_{i_1,\cdots,i_k | \Phi_l^d}} \big[ \text{log} P(i_1 | \Phi_l^d) \big] \\
&= -\displaystyle{\mathop{\mathbb{E}}_{i_1 | \Phi_l^d}} \big[ \text{log} P(i_1 | \Phi_l^d) \big] = \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big]
\end{align*}
\section{Appendix C}
\begin{breakablealgorithm}
\caption{Training \system{LaMGD}} \label{alg:mdmt}
\begin{algorithmic}[1]
\REQUIRE {
  \begin{itemize}
    \setlength{\itemsep}{-1pt}    \setlength{\parsep}{0pt}
  \item $n_d$ corpora $C^d, d \in [1, \dots,n_d]$ for $n_d$ domains equipped by an empirical distribution $D_d(x)$
  \item number of retained groups: k, number of groups: $n_p$
  \item $i=0$
 \end{itemize}}
\REPEAT 
\STATE{Pick a batch from domain $d$}
\STATE{Sampling $\forall p,  g_l^d(p) \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1)$}
\STATE{Solve the equation
\begin{align*}
\displaystyle{\mathop{\sum}_{i=1}^{n_p}} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) = k
\end{align*}
using binary search}
\STATE{Computing mask of each layer
\begin{align*}
\hat{m}_l^d(\tau) &= \sigma(\frac{g_l^d + \Phi_l^d + \bar{\nu}}{\tau})
\end{align*}}
\STATE{Applying computed masks to their corresponding layer
\begin{align*}
  \forall l \in [0,\cdots, L-1]: \tilde{h}^l &= h^l \odot r_l^d ,\\
  h^{l+1} &= \operatorname{LAYER}^{l+1}(\tilde{h}^l) ,
\end{align*}
}
\STATE{Compute gradient of training loss over the underlying Transformer
$$\Delta_{\theta} = \frac{\partial L}{\partial \theta}$$}
\STATE{Compute gradient over the $\operatorname{Soft-Top-K}$ masks
$$\frac{\partial D}{\partial \hat{m}_l^d(\tau)}$$}
\STATE{Compute implicit gradient of the $\operatorname{Soft-Top-K}$ masks over $\Phi^d$
$$- \big(\frac{\partial g}{\partial \bar{\nu}}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d}$$}
\STATE{Compute the gradient the training over $\Phi$
$$ - \frac{\partial D}{\partial \hat{m}_l^d(\tau)} \times \big(\frac{\partial g}{\partial \bar{\nu}}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d} + \frac{\partial \mathbb{H} \big[ \operatorname{softmax}(\Phi_l^d)\big] }{\partial \Phi_l^d} $$}
\STATE{Update $\theta$ and $\Phi^d$ with their gradients}
\STATE{$i=i+1$}
\UNTIL{$i> iter\_num$}
\end{algorithmic}
\end{breakablealgorithm}
\section{Appendix D}
\begin{figure*}[h!]
\includegraphics[width=0.7\textwidth]{multi_domain}
\caption{Visualization of domains according to their dropouts (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
\label{fig:domain}
\end{figure*}
\end{document}

>>>>

\bibliography{bibliography}
\bibliographystyle{acl_natbib}
\appendix
\section{Appendix A}
\label{appendix:a}
This section explains how to compute $\hat{m}_l^d(\tau)$ by solving the optimization problem \eqref{eq:soft-top-k} and then how to compute the gradients $\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d}$.

First, to solve \eqref{eq:soft-top-k} we follow the same approach as in \cite{amos19lml,amos20differential} by applying the Karush–Kuhn–Tucker (KKT) conditions to~\eqref{eq:soft-top-k}. The solution of \eqref{eq:soft-top-k} will have the following form:
\begin{align}
\hat{m}_l^d(\tau) &= \sigma(\frac{g_l^d + \Phi_l^d + \bar{\nu}}{\tau}) \label{eq:soft-m}
\end{align}
in which $\sigma(.)$ is the sigmoid function and $\bar{\nu}$ is the solution of the following equation:
\begin{align}
\displaystyle{\mathop{\sum}_{i=1}^{n_p}} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) = k \label{eq:nu}
\end{align}

Because sigmoid is monotonous, equation~\eqref{eq:nu} has a unique solution. Furthermore,  because of the smoothness of $g(\nu,\Phi_l^d) = \displaystyle{\mathop{\sum}_{i=1}^{n_p}} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau})$ w.r.t $\nu$ and $\Phi_l^d$, we can perform the implicit differentiation of its solution $\bar{\nu}$ w.r.t $\Phi_l^d$ as below, even though the solution of \eqref{eq:nu} does not have an explicit form.

@@

Because the differentiation of sigmoid has exact forms, $\frac{\partial g}{\partial \nu}$ and $\frac{\partial g}{\partial \Phi_l^d}$ also have exact form. Therefore, we do not need autograd to compute the implicit gradient $\frac{\partial \nu}{\partial \Phi_l^d}$. The gradient of $\hat{m}_l^d(\tau)$ w.r.t $\Phi_l^d$ is computed as follows:
\begin{align}
\frac{\partial \hat{m}_l^d(\tau)}{\partial \Phi_l^d} = \frac{\partial \hat{m}_l^d(\tau)}{\partial \nu} \times \frac{\partial \nu}{\partial \Phi_l^d}
\end{align}

In our algorithm, we solve \eqref{eq:nu} by binary search. The convergence of binary search is extremely fast and assured by the monotonicity of $g(\nu,\Phi_l^d)$. In our experiments, we set the search range to $[-100,100]$.

Finally, we need prove that $\lim_{\tau \rightarrow 0}\hat{m}_l^d(\tau) = \tilde{m}_l^d$. We assume $g_l^d(i_1) + \Phi_l^d(i_1) > g_l^d(i_2) + \Phi_l^d(i_2) > \cdots > g_l^d(i_{n_p}) + \Phi_l^d(i_{n_p})$.

Because:
\begin{align*}
\hspace{-3.em}
\lim_{\tau \rightarrow 0} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) = \begin{cases}
      1, & \text{if}\ \tau > - (g_l^d(i) + \Phi_l^d(i)), \\
      0, & \text{if}\ \tau < - (g_l^d(i) + \Phi_l^d(i)), \\
      \frac{1}{2} &\text{otherwise}\\
    \end{cases}
\end{align*}

@@

\begin{align*}
\displaystyle{\mathop{\sum}_{i=1}^{n_p}} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) = k
\end{align*}
there exist $\epsilon$ such that $\forall \tau < \epsilon$, the solution $\bar{\nu}$ of \eqref{eq:nu} satisfies $- (g_l^d(i_{k+1}) + \Phi_l^d(i_{k+1})) > \bar{\nu} > - (g_l^d(i_k) + \Phi_l^d(i_k))$. Furthermore, because sigmoid is monotonous,

@@

By taking the limit on both sides, we get the following results:
\begin{align*}
\lim_{\tau \rightarrow 0} \hat{m}_l^d(\tau)(i_u) = \begin{cases}
      1, & \text{if}\ u > k \\
      0, & \text{if}\ u < k \\
    \end{cases}
\end{align*}

And, because $ \displaystyle{\mathop{\sum}_{u=1}^{n_p}}\hat{m}_l^d(\tau)(i_u) = k$, by taking the limit on both sides, we will have $\lim_{\tau \rightarrow 0} \hat{m}_l^d(\tau)(i_k) = 1$. Finally, we have

@@

\section{Appendix B}
\label{appendix:b}
In this section, we give a simple proof of inequality~\eqref{eq:entropy}. In fact, we only need to prove $\mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] \geqslant \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big]$. The proof is as follows:
\begin{align*}
&\mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] = - \displaystyle{\mathop{\mathbb{E}}_{i_1,\cdots,i_k | \Phi_l^d}} \big[ \text{log}P(i_1,\cdots,i_k | \Phi_l^d) \big] \\
&= -\displaystyle{\mathop{\mathbb{E}}_{i_1,\cdots,i_k | \Phi_l^d}} \big[ \displaystyle{\sum_{j=2}^k}\text{log}P(i_j | i_1,\cdots,j_{j-1},\Phi_l^d) +  \text{log} P(i_1 | \Phi_l^d) \big] \\
&\geqslant -\displaystyle{\mathop{\mathbb{E}}_{i_1,\cdots,i_k | \Phi_l^d}} \big[ \text{log} P(i_1 | \Phi_l^d) \big] \\
&= -\displaystyle{\mathop{\mathbb{E}}_{i_1 | \Phi_l^d}} \big[ \text{log} P(i_1 | \Phi_l^d) \big] = \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big]
\end{align*}
\section{Appendix C}
\begin{breakablealgorithm}
\caption{Training \system{LaMGD}} \label{alg:mdmt}
\begin{algorithmic}[1]
  \REQUIRE {%    
  \begin{itemize}
    \setlength{\itemsep}{-1pt}    \setlength{\parsep}{0pt}
  \item[] 
  \item $n_d$ corpora $C^d, d \in [1, \dots,n_d]$ for $n_d$ domains equiped by an empirical distribution $D_d(x)$
  \item number of groups: $n_p$; number of retained groups: $k$ 
  \item $i=0$
 \end{itemize}}
\REPEAT 
\STATE{Pick a batch from domain $d$}
\STATE{Sample $\forall l, \forall p:  g_l^d(p) \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1)$}
\STATE{Solve the equation $\forall l$
\begin{align*}
\displaystyle{\mathop{\sum}_{i=1}^{n_p}} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) = k
\end{align*}
using binary search}
\STATE{Compute mask of each layer
\begin{align*}
\forall l, \hat{m}_l^d(\tau) &= \sigma(\frac{g_l^d + \Phi_l^d + \bar{\nu}}{\tau})
\end{align*}}
\STATE{Apply masks to their corresponding layer
\begin{align*}
  \forall l \in [0,\cdots, L-1]: \tilde{h}^l &= h^l \odot r_l^d ,\\
  h^{l+1} &= \operatorname{LAYER}^{l+1}(\tilde{h}^l) ,
\end{align*}
}
\STATE{Compute gradient of training loss over the underlying Transformer
$$\Delta_{\theta} = \frac{\partial L}{\partial \theta}$$}
\STATE{Compute gradient over the $\operatorname{Soft-Top-K}$ masks
$$\frac{\partial D}{\partial \hat{m}_l^d(\tau)}$$}
\STATE{Compute implicit gradient of the $\operatorname{Soft-Top-K}$ masks over $\Phi^d$
$$- \big(\frac{\partial g}{\partial \bar{\nu}}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d}$$}
\STATE{Compute the gradient the training over $\Phi$
$$ - \frac{\partial D}{\partial \hat{m}_l^d(\tau)} \times \big(\frac{\partial g}{\partial \bar{\nu}}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d} + \frac{\partial \mathbb{H} \big[ \operatorname{softmax}(\Phi_l^d)\big] }{\partial \Phi_l^d} $$}
\STATE{Update $\theta$ and $\Phi^d$ with their gradients}
\STATE{$i=i+1$}
\UNTIL{$i> iter\_num$}
\end{algorithmic}
\end{breakablealgorithm}
\section{Appendix D}\fyTodo{fix this}
\begin{figure*}[h!]
\includegraphics[width=0.7\textwidth]{multi_domain}
\caption{Visualization of domains according to their dropouts (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
\label{fig:domain}
\end{figure*}
\end{document}
